A C-to-RTL flow as an energy efficient alternative to embedded processors in digital systems

### 2010-05-27 Documents/Publications/dsd2010/paper.tex
Modified paper.tex to add a line in the introduction. No other changes
Note: have also prepared paper.ps and sent to IEEE
pdfExpress to convert ot pdf.  If this tex file is
final, we can submit the pdf file generated by
pdfExpress.

We provide comparisons between
the hardware generated by our high level synthesis flow and  
processor implementations of an algorithm described in C.  
Note that it is not entirely obvious that algorithm-specific hardware 
generated by a high-level synthesis flow will be superior
to embedded processor implementations of the same algorithm.  
One can argue that the algorithm-specific hardware
can exploit the characteristics of the algorithm to a greater
degree than a general purpose processor.  
On the other hand, when a C/C++ program is compiled to
a processor, the platform (that is, the processor) is usually a
heavily optimized one in which each component has been tweaked to a
high level, whereas the algorithm-specific
hardware is generated by automated tools whose optimization abilities
will most probably not match those of a human designer.   
Concrete data which throws light on these aspects is valuable.

@@

In order to characterize the energy efficiency of the resulting
circuits, we select a wide range of applications (from stream ciphers
to databases and linear algebra) and run them through our C-to-RTL
flow. The resulting RTL is then mapped to an ASIC using industry
standard automatic synthesis tools. This ASIC is characterized for
power, delay, area and energy, and the results are compared with those
obtained when the same program is mapped to a low power processor.
These results indicate that even in its current state, our high-level
synthesis flow is an energy competitive alternative to the use of
embedded processors in the design of complex systems.

@@

In Table \ref{table:Ratios}, we tabulate the ratios of
the performance parameters area, frequency, delay, power
and energy for the processor $P$ relative to the scaled values
for the AHIR circuits assuming the scaling rules shown above.
From the table, we see that across all applications, the energy consumed
by AHIR circuits is considerably lower than the
processor (the energy per task for the processor is
between $4.7X$ and $116.6X$ that of the corresponding AHIR circuit).  
The improvement is highest in the case of
bit-level manipulation applications such as A5/1, and
lowest in the case of random ``pointer-chasing'' code.
The delays in completing the tasks are lower 
for the processor $P$ in almost all the cases.  This
is mainly due to the fact that currently, the operators used in 
the AHIR generated RTL are not pipelined or optimized in any
manner during the logic synthesis and physical
design. This limits the operating clock frequency of the circuits
synthesized from the AHIR generated RTL.

>>>>

We provide comparisons between
the hardware generated by our high level synthesis flow and  
processor implementations of an algorithm described in C.  
Note that it is not entirely obvious that algorithm-specific hardware 
generated by a high-level synthesis flow will be superior
to embedded processor implementations of the same algorithm.  
One can argue that the algorithm-specific hardware
can exploit the characteristics of the algorithm to a greater
degree than a general purpose processor.  
On the other hand, when a C/C++ program is compiled to
a processor, the platform (that is, the processor) is usually a
heavily optimized circuit in which each functional
element has been tweaked to a
high level, whereas the algorithm-specific
circuit is generated by automated tools whose optimization abilities
will most probably not match those of a human designer.   
Finally, since it is not possible to express parallelism
directly in a C program (although a compiler can infer it),
it is not too clear whether a high-level synthesis system can
identify enough parallelism to make the resulting hardware competitive
with a heavily optimized processor.  Concrete data which throws light on these aspects is valuable.

@@

In order to characterize the energy efficiency of the resulting
circuits, we select a wide range of applications (from stream ciphers
to databases and linear algebra) and run them through our C-to-RTL
flow. The resulting RTL is then mapped to an ASIC using industry
standard automatic synthesis tools. This ASIC is characterized for
power, delay, area and energy, and the results are compared with those
obtained when the same program is mapped to a low power processor.
The results indicate that even in its current state, our high-level
synthesis flow is an energy competitive alternative to the use of
embedded processors in the design of complex systems.

@@

In Table \ref{table:Ratios}, we tabulate the ratios of
the performance parameters area, frequency, delay, power
and energy for the processor $P$ relative to the scaled values
for the AHIR circuits assuming the scaling rules shown above.
From the table, we see that across all applications, the energy consumed
by AHIR circuits is considerably lower than the
processor (the energy per task for the processor is
between $4.7X$ and $116.6X$ that of the corresponding AHIR circuit).  
The improvement is highest in the case of
bit-level manipulation applications such as A5/1, and
lowest in the case of random ``pointer-chasing'' code.  Note
that the improvement in the
case of  floating point programs (FFT and LPK) is also
substantial.
The delays in completing the tasks are lower 
for the processor $P$ in almost all the cases.  This
is mainly due to the fact that currently, the operators used in 
the AHIR generated RTL are not pipelined or optimized in any
manner during the logic synthesis and physical
design. This limits the operating clock frequency of the circuits
synthesized from the AHIR generated RTL.
